CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-07 codecheck-paper.tex
edits to principles section

The codechecker
community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
currently has over 20 individuals who volunteered in the last 12
years.  These volunteer's motivations, mentioned in the registration
information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
include: supporting reproducible research and Open Science, improve
coding skills, gaining experience in helping scientists with their
code, encouraging a sharing culture, and learning from other people's
mistakes; many are also motivated simply by curiosity.  We see
benefits to an open shared list of codecheckers across journals rather
than a private in-house group, as this may allow for better matches
regarding expertise and more even sharing of the workload.  Overall,
this community will aim to ensure that CODECHECK is a viable option
for independent no-cost Open Access journals.

@@

The example workflow and variations outlined above reflect our current
views on how a workflow execution and code checking should be performed as part of peer review.
They are not set in stone, but we do believe the following core principles underpin our CODECHECK process:

\textbf{1. Codecheckers record but don't investigate or fix.} \\
The codechecker follows the instructions provided by the author in running
the code. If any steps are unclear, or if code does not run correctly,
the codechecker records this fact and then communicates this
information to the author. We believe that the job of the codechecker is
not to fix all these problems but simply to report them and await a
fix from the author.
The level of documentation that is needed for third parties
to reproduce a workflow is extremely hard to get right, and too often this 
uncertainty leads researchers to give up and not document it at all.
The conversation with a codechecker fixes this problem.

\textbf{2. Communication between humans is key.} \\ 
Some code may
work without any interaction (e.g., CODECHECK certificate \texttt{2020-013}
\cite{cert-2020-013}) but more often there are hidden
dependencies that need to be found or software installations that need to be adjusted for a particular system.
Allowing the (human) codechecker to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community building as well as obscure some of the lessons
learned for both authors and codecheckers.

\textbf{3. Credit is given to codecheckers.} \\ 
The value of performing a
CODECHECK is comparable to that of a scientific peer review, and
it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e.,~with a DOI, as conducted in the community CODECHECK
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases.

\textbf{4. Workflows must be auditable.} \\ 
This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this ``preproducibility''
and the ICERM report \cite{stodden_setting_2013} defines the level
``Auditable Research'' similarly.
Every community may establish their own good practices or adapt generic
concepts and practical tools, such as publishing all building blocks
of science in a research compendium
\footnote{\url{https://research-compendium.science/}}.
A completed check means that code could be executed at least once without
critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a ``one~click'' step, but achieving this requires particular 
skills and a sufficient level of documentation for 
third parties. Furthermore, automation may lead to people gaming the system
or reliance on technology, which can often hide important details.
All such aspects can reduce the understandability of the material, so we estimate our approach to codechecking, done without automation and with open human communication, to be the best way to ensure long-term transparency and usefulness.
We acknowledge that others have argued in favour of bitwise reproducibility
because, in the long run, it can be automated\footnote{Konrad Hinsen on 
Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}},
but until then we need CODECHECK's approach.

\textbf{5. Open by default and transitional by disposition.} \\ 
By default, unless there are strong
reasons to the contrary (e.g.,~sensitive data on human subjects), all code and
data, both from author and codechecker, will be made freely available, 
i.e.,~under open licenses, before or when the certificate is published.
Note that openness is not required for the paper itself, as to accommodate 
journals in their transition to sustainable Open Access models.
The code and data publication should follow community good practices.
By disposition, a CODECHECK workflow should be designed (a) with a recognition of
shortcomings in asserting reproducibility of submissions and (b)
with an understanding of the
complex situation that led to this. However, establishing a CODECHECK workflow
should go hand in hand with initiatives aiming to improve education
around reproducibility of methods and communication surrounding comprehensive 
computational workflows.
The long term goal is for a specific community or
journal's target audience to reach a state where the level of verification 
ensured by codecheckers becomes part of peer review and a regular
merit of all accepted papers.

@@

>>>>

The codechecker
community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
has over 20 volunteers who signed up in the last year.  Their
motivations, mentioned in the registration
information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
include: supporting reproducible research and Open Science, improve
coding skills, gaining experience in helping scientists with their
code, encouraging a sharing culture, and learning from other people's
mistakes; many are also motivated simply by curiosity.  We see
benefits to an open shared list of codecheckers across journals rather
than a private in-house group, as this may allow for better matches
regarding expertise and more even sharing of the workload.  Overall,
this community will aim to ensure that CODECHECK is a viable option
for independent no-cost Open Access journals.

@@

The workflow and variations outlined describe our current
views on how code could be checked.  They are not immutable, but
we believe the following core principles underpin our CODECHECK
process:

\textbf{1. Codecheckers record but don't investigate or fix.} \\
The codechecker follows the author's instructions to run the code. If
instructions are unclear, or if code does not run, the codechecker
tells the author. We believe that the job of the codechecker is not to
fix all these problems but simply to report them and await a fix from
the author.  The level of documentation that is needed for third
parties to reproduce a workflow is hard to get right, and
too often this uncertainty leads researchers to give up and not
document it at all.  The conversation with a codechecker fixes this
problem.

\textbf{2. Communication between humans is key.} \\ 
Some code may
work without any interaction (e.g., CODECHECK certificate \texttt{2020-013}
\cite{cert-2020-013}) but more often there are hidden
dependencies or installations that need adjusting for a particular system.
Allowing the (human) codechecker to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would introduce delays and
reduce the chances for community building.

\textbf{3. Credit is given to codecheckers.} \\
The value of performing a CODECHECK is comparable to that of a peer
review, and it may also require a similar amount of time. Therefore,
the codechecker's activity should be recorded, ideally in the
published paper.  The public record can be realised by publishing the
certificate in a citable form (i.e.,~with a DOI, as conducted in the
community CODECHECK process), by listing codecheckers on the journal's
website or, ideally, by publishing the checks alongside peer review
activities in public databases.

\textbf{4. Workflows must be auditable.} \\
The codechecker should have sufficient material to validate the
workflow outputs submitted by the
authors. Stark~\cite{stark_before_2018} calls this
``preproducibility'' and the ICERM report \cite{stodden_setting_2013}
defines the level ``Auditable Research'' similarly.  Every community
may establish their own good practices or adapt generic concepts and
practical tools, such as publishing all building blocks of science in
a research compendium
\footnote{\url{https://research-compendium.science/}}.  A completed
check means that code could be executed at least once using the
provided instructions, and, therefore, all code and data was given and
could be investigated more deeply or extended in the future.  Ideally,
this is a ``one~click'' step, but achieving this requires particular
skills and a sufficient level of documentation for third
parties. Furthermore, automation may lead to people gaming the system
or reliance on technology, which can often hide important details.
All such aspects can reduce the understandability of the material, so
we estimate our approach to codechecking, done without automation and
with open human communication, to be a simple way to ensure long-term
transparency and usefulness.  We acknowledge that others have argued
in favour of bitwise reproducibility because, in the long run, it can
be automated\footnote{Konrad Hinsen on Twitter:
  \url{https://twitter.com/khinsen/status/1242842759733665799}}, but
until then we need CODECHECK's approach.

\textbf{5. Open by default and transitional by disposition.} \\
By default, unless there are strong reasons to the contrary
(e.g.,~sensitive data on human subjects), all code and data, both from
author and codechecker, will be made freely available before or when
the certificate is published.  Openness is not required for the paper
itself, to accommodate journals in their transition to sustainable
Open Access models.  The code and data publication should follow
community good practices.  Ultimately we may find that CODECHECKS are
subsumed within peer review.
%% By disposition, a CODECHECK workflow should
%% be designed (a) recognising shortcomings in asserting reproducibility
%% of submissions and (b) with an understanding of the complex situation
%% that led to this. However, establishing a CODECHECK workflow should
%% mirror initiatives to improve education around reproducibility of
%% methods.  The long term goal is for a specific community to reach a
%% state where the level of verification ensured by codecheckers becomes
%% part of peer review and a regular merit of all accepted papers.

@@

% Local Variables:
% TeX-engine: xetex
% End:
